---
title: "Hello World!"
output: 
  distill::distill_article:
    toc: FALSE
site: distill::distill_website
preview: photos/uzumaki.png
css: [style.css]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
## Welcome to One Carat Blog!
 
```{r, echo=FALSE, layout="l-screen", fig.height=3}
library(r2d3)
r2d3(d3_version = 5, script = "voronoi/voronoi.js")
```

Welcome to my blog! My name is Jewel Johnson and I am the founder of this blog. Here you will find cool science articles and data visualizations. If you want to learn about R programming, check out the R tutorial tab on top. For blog posts, check the blog section.

So give this blog a try and let me know how it is! Thanks for visiting and have a good day!

<center>Visitors<center>
<center><div id="sfcxqmebarkplgkzwtm8bz2lp8pnnllbeal"></div>
<script type="text/javascript" src="https://counter7.stat.ovh/private/counter.js?c=xqmebarkplgkzwtm8bz2lp8pnnllbeal&down=async" async></script>
<noscript><a href="https://www.freecounterstat.com" title="website counter code"><img src="https://counter7.stat.ovh/private/freecounterstat.php?c=xqmebarkplgkzwtm8bz2lp8pnnllbeal" border="0" title="website counter code" alt="website counter code"></a></noscript>
</div><center>


