---
title: " "
description: |
  Learn how to 
author:
  - name: Jewel Johnson 
date: 12-23-2021
base_url: https://jeweljohnsonj.github.io/jeweljohnson.github.io/
output: 
  distill::distill_article:
    toc_depth: 3
preview: photos/dplyr.png
---
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(palmerpenguins)
library(rmarkdown)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

```{r, xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clone fa-2x\" style=\"color: #301e64\"></i>",
    success_text = "<i class=\"fa fa-check fa-2x\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times fa-2x\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

---
# xaringanExtra package will help us to have inbuilt tabs insdie the article
---

```{r panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

---
# loading javascipt file which helps in folding outputs similar to code folding
---
<script src="js/output_folding.js"></script>


---
#start editing from here
---

## Heading 1



## Summary


---
# buttons for navigation
---

<br>  

<a href="project5.html" class="btn button_round" style="float: left;">Previous chapter:<br> 2: Data manipulation using dplyr (part 1)</a>

## References



## Last updated on {.appendix}

```{r, echo=FALSE}
Sys.time()
```

---
# social media share buttons
---

<!-- adding share buttons on the right side of the page -->
<!-- AddToAny BEGIN -->
<div class="a2a_kit a2a_kit_size_32 a2a_floating_style a2a_vertical_style" style="right:0px; top:150px; data-a2a-url="https://jeweljohnsonj.github.io/jeweljohnson.github.io/" data-a2a-title="One-carat Blog">
<a class="a2a_button_twitter"></a>
<a class="a2a_button_whatsapp"></a>
<a class="a2a_button_telegram"></a>
<a class="a2a_button_google_gmail"></a>
<a class="a2a_button_pinterest"></a>
<a class="a2a_button_reddit"></a>
<a class="a2a_button_facebook"></a>
<a class="a2a_button_facebook_messenger"></a>
</div>
<script>
var a2a_config = a2a_config || {};
a2a_config.onclick = 1;
</script>
<script async src="https://static.addtoany.com/menu/page.js"></script>
<!-- AddToAny END -->

